LaMachine: A meta-distribution for NLP software

### 2019-03-29 docs/lamachine_clarin2018.tex
further work on paper

The CLARIN Switchboard \cite{zinn2016clarin} also offers access to NLP tools at a different level as it provides a webinterface where users can upload their texts for NLP processing, proving very easy acces. Lamachine is most often used by data scientists and reseachers who do have programming experience and want to work with complex NLP software.

Comparisons could be drawn between LaMachine as a kind of Linux distribution and other Linux distributions. That,
however, would not be appropriate as we characterise LaMachine more as a meta-distribution which simply builds on the foundation
of several existing major Linux distributions. As a meta-distribution that runs in a variety of flavours and on a
variety of platforms, it offers a great deal of flexibility that normal distributions do not have. At the same time, the scope of
LaMachine is more narrowly defined than that of a generic Linux distribution, including even specialised Linux
distributions that focus on scientific computing, such as Fedora
Scientific\footnote{\url{https://labs.fedoraproject.org/en/scientific/}}.

@@

Virtual Research Environments (VRE) such as the Dutch CLARIAH WP3 VRE project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}} and the Spanish VRE project \cite{gonzalez2017evi} aim to develop more high level tool environments with graphical interfaces to accomodate the reseacher. Lamachine is currently developed in close collaboration with the Dutch CLARIAH WP3 VRE project.

@@

LaMachine is most often used by a researcher or data scientist who installs LaMachine on his own computer (locally, in a
local environment or as a Docker container or VM). Here we discuss two case studies where LaMachine was used under
circumstances that go beyond this most common usage of LaMachine: using LaMachine on a secure server without internet
connection, and using LaMachine to provide a work environment for a group of researchers. These case studies illustrate the advantages and limitations of Lamachine in practice.

@@

We used a virtual machine (VM) image of LaMachine and we ran this 64-bits Linux-based VM inside another
VM with Windows Server 2012, provided to us by the health inspectorate for this project, in which we did have administrative
rights but no internet access. In terms of hardware we ran on a machine with 8 cores and  32GB internal memory available. Note that as we had no internet access available on the working server, we prepared a custom-made VM image of Lamachine that included an additional editor and some extra Python libraries.
%using VirtualBox and Vagrnt on the Windows VM
LaMachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use
the Jupyter Notebook environment.

@@

Most users of Lamachine are installing the software distribution in their own machine, either locally, via Docker or VM. However, we described to use cases in section \ref{sec:case} where we explore the possibilities under more constrained and demanding circumstances.
Such case studies contribute to thorough testing and running of LaMachine in less ideal circumstances such as nested VM constructions and offline usage.

>>>>

Within the scope of CLARIN, one might be inclined to draw comparisons with the CLARIN Language Resource Switchboard
\cite{zinn2016clarin}, which also offers access to NLP tools through a web interface where users can upload their texts
for NLP processing, providing easy access and automated tool discovery.  This, however, is notably different from what
LaMachine is. The CLARIN switchboard is not a software distribution, it links to externally hosted applications; and
LaMachine does currently not provide any switchboard functionality, but merely offers a portal interface. Moreover,
LaMachine is used by data scientists and reseachers who do have programming experience and want to work with complex NLP
software, on their own machines and unencumbered by high-level interfaces if they so desire.

There are also CLARIN initiatives that provide not just a switchboard, but a more elaborate high-level virtual research
environment (VRE) with graphical interfaces to accommodate the researcher and allow execution and visualisation of
various NLP pipeline. The Dutch CLARIAH WP3 project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}}
is such an initiative. LaMachine is currently developed in
close collaboration with the Dutch CLARIAH WP3 VRE project, in which we act as a technology supplier.

Comparisons could be drawn between LaMachine as a kind of Linux distribution and other Linux distributions. That too,
however, would not be appropriate as we characterise LaMachine more as a meta-distribution which simply builds on the foundation
of several existing major Linux distributions. As a meta-distribution that runs in a variety of flavours and on a
variety of platforms, it offers a great deal of flexibility that normal distributions do not have. At the same time, the scope of
LaMachine is more narrowly defined than that of a generic Linux distribution, including even specialised Linux
distributions that focus on scientific computing, such as Fedora
Scientific\footnote{\url{https://labs.fedoraproject.org/en/scientific/}}.

@@

@@

LaMachine is most often used by a researcher or data scientist who installs LaMachine on his own computer (locally, in a
local environment or as a Docker container or VM). Here we discuss two case studies where LaMachine was used under
circumstances that go beyond this most common usage of LaMachine: using LaMachine on a secure server without internet
connection, and using LaMachine to provide a work environment for a group of researchers. These case studies illustrate
the advantages and limitations of LaMachine in practice.

@@

We used a virtual machine (VM) image of LaMachine and we ran this 64-bits Linux-based VM inside another
VM with Windows Server 2012, provided to us by the health inspectorate for this project, in which we did have administrative
rights but no internet access. In terms of hardware we ran on a machine with 8 cores and  32GB internal memory available. Note that as we had no internet access available on the working server, we prepared a custom-made VM image of LaMachine that included an additional editor and some extra Python libraries.
%using VirtualBox and Vagrnt on the Windows VM
LaMachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use
the Jupyter Notebook environment.

@@

Most users of LaMachine are installing the software distribution in their own machine, either locally, via Docker or VM. However, we described to use cases in section \ref{sec:case} where we explore the possibilities under more constrained and demanding circumstances.
Such case studies contribute to thorough testing and running of LaMachine in less ideal circumstances such as nested VM constructions and offline usage.
